\chapter{Method}

\textcolor{red}{Skal jeg få inn litt om etikk her? Er jo litt fare at jeg er både utvikler i offentlig sektor og forsker. Kan gi "validity" problems. Er også problemer med at de andre ansatte kan lettere gi med sensitive dokumenter/opplysninger, som ikke er egnet å publiseres (større ansvar for det etiske pga min posisjon)}

%\section{Research Questions} 
%This study has the main objective of understanding how the different agencies in the Norwegian public sector assure quality in their software. The study also have the objective of understanding how well the software quality assurance in the Norwegian public sector compares with what is recommended by the relevant literature shown in \autoref{sec:related_research}. This leading to the following research questions:

%\begin{itemize}
%    \item \textbf{RQ1:} How do the different Norwegian public sector's assure quality of their software?
%    \item \textbf{RQ2:} How does the process of software quality assurance in the Norwegian public sector compare to what is stated in relevant literature?
%\end{itemize}

%\section{Method}
As stated, the objective of this study is to understand what the people with high influence or ownership over technology in the Norwegian public sector considers important in regards to assuring quality in its software and what factors might influence their view on this. Due to this being a broader theme or phenomena \cite{bjo_2022} without a clear answer as shown in \autoref{sec:related_research}, it is decided that a case study with the context of the Norwegian public sector is used as an appropriate method. The case study uses the experience of the researcher as a software developer in the Norwegian public sector and interview's as data generation methods.

\section{Selection of method}
There are several reasons for why case study is chosen as the method for this study \cite{bjo_2022}, where the main reason is that software quality assurance in the Norwegian has little research. As well as the term "software quality assurance" can be interpreted in multiple ways with different points of views. A case study therefor allows the researcher to gain a deeper understanding of the topic from multiple sources, both qualitative and quantitative, which gives the researcher a larger range of data to be interpreted \cite{bjo_2022}.

\section{Case Study}

\subsection{Type of case study}
\paragraph{Exploratory study}
As there is little research on the research topic in this study, and what research does exist is mostly broad and qualitative, it is decided that an exploratory study is used to gain an deeper understanding in the research topic \cite{bjo_2022}.

\paragraph{Short term, contemporary study} 
The case study is a short term, contemporary study as the researcher is interested in what is the understanding of software quality assurance in the Norwegian public sector at the time of this study \cite{bjo_2022}. It is also decided due to the limited time resources in the production of the study, hence a longitudinal study would not be possible to produce.

\subsection{Selection of case}
There are several for why software quality assurance in the Norwegian public sector is subject to a case study

\paragraph{Typical instance}
Software quality assurance is important for other areas than just in the Norwegian public sector. The results of this case study could be applied to other nations public sectors, or other organisations in different private sectors.

\paragraph{Convenience}
As the researcher works as a software developer in the Norwegian public sector, it is a convenient case to study. The researcher already has contact with potential participants for the study and in-direct contacts with other potential participants in the Norwegian public sector.

\paragraph{Unique opportunity}
The researchers unique position as both a researcher and software developer in the Norwegian public sector gives an unique opportunity for the chosen case. The researcher is able to bring first hand accounts with own experience while also conducting research which could support or contradict those experiences.


\subsection{Data generation methods}

\subsubsection{Interviews}

\textcolor{red}{Huske å få inn her at jeg vil også ha "expert interviews" fra forskere i feltet}

\paragraph{Semi-structured}

\paragraph{Recording}
\begin{itemize}
    \item Field notes
    \item Audio recording
\end{itemize}


\begin{comment}
In order to gain a understanding in how the Norwegian public sector assures quality in its software, it was decided to do a survey involving different departments in the Norwegian public sector. The method for the survey follows the recommendations of the book "Researching Information Systems and Computing", and includes the following steps: (I) Data requirements, (II) Data generation method, (III) Sampling frame, (IV) Sampling technique, (V) Response rate and non-respondents, and (VI) Sample size \cite{bjo_2022}.

\subsection{Data requirements}
The data needed to be collected is both directly and in-directly topic-related. The directly topic-related data requirements relates to specifics in the research questions, such as what methods are used for assuring quality in software. The in-direct topic-related data is used for connecting the directly-topic related data \cite{bjo_2022}, i.e. connecting software quality assurance to the context of the Norwegian public sector.

%Directly topic-related -> developers, designers etc are directly involved in assuring the quality of software

\subsubsection{Directly topic-related}
The following data is collected that is directly topic-related:
\begin{itemize}
    \item Non-functional requirements.
    \item Software quality assurance methods.
    \item Confidence in software quality assurance method.
\end{itemize}

\subsubsection{In-Directly topic-related}
The following data is collected that is in-directly topic-related:
\begin{itemize}
    \item Agency
    \item Role/technical responsibility
    \item Permanent employee or consultant
\end{itemize}

\subsection{Data generation method}

\subsubsection{Questionnaire}
A self-administered questionnaire is used to gain insight into the surveys research questions. As recommended  \cite{bjo_2022}, there are several reasons for using a self-administered questionnaire:

\begin{itemize}
    \item The survey intends to gather brief and uncontroversial data from a large number of people
    \item The data in the survey needs to be standardised
    \item Geographical location between potential respondents and researchers makes personal interviews unpractical
    \item Researcher does not have time for personal interviews, but has time to wait between distributing the questionnaire and getting responses
    \item In-experience by the researcher can lead to cognitive bias in the data if questionnaire was researcher-administered
\end{itemize}


\subsubsection{Structure of questionnaire}
\textcolor{red}{Når den er mer spikra, skrive om den her
\begin{enumerate}
    \item Hvilken etat tilhører du?
    \item Hvilken rolle har du? (Skal utvikler splittes opp i separate roller? Frontend, backend osv..)
    \item Er du internt ansatt eller innleid (konsulent)?
    \item Hvilke av disse har du fokus på i rollen din? (Liste med non-functionals requirements)
    \item Hvilke metoder bruker du for å sikre at disse er av kvalitet? (Liste med metoder? Fritekst?)
    \item På en skala fra 1-5, hvordan synes du disse funker? (Tall skala eller "ord" skala).
\end{enumerate}
}

\subsection{Sampling frame}
The following agencies in the Norwegian public sector will be used as the sampling frame for the survey, mainly targeting each agency's IT-department.

\begin{table}[H]
\begin{tabular}{ll}
\hline
Department & Agency \\ \hline
           &        \\
           &        \\
           &        \\
           &        \\ \hline
\end{tabular}%
\centering
\caption{Sampling frame of Norwegian public sector agencies included in the survey}
\label{tab:sample_frame}
\end{table}

\subsection{Sampling technique}
The sampling technique for the survey is primarily probabilistic, as it is believed that the respondents of the survey is representative to the population of the area being researched.

Cluster sampling

\subsection{Response rate}

\subsection{Non-respondents}


\subsection{Data synthesis}
\end{comment}